Declarative deployment reconfiguration of pulverised collective adaptive systems at runtime \thanks{Identify applicable funding agency here. If none, delete this.}

### 2023-05-15 paper-2023-acsos-pulverization.tex
add some details in sections ii and iii

\acrodef{dsl}[DSL]{domain-specific language}
\acrodef{iot}[IoT]{Internet of Things}

@@

Recent research trends in topics related to the Internet of Things promote the idea of a cloud-edge continuum~\cite{DBLP:journals/iot/BittencourtISFM18},
an infrastructure composed of heterogeneous devices ranging from large and powerful cloud servers to small connected things
on which software computes in a distributed fashion,
optimising for performance and resource utilisation.
%
Exploiting this intriguing infrastructure, however, poses new challenges to the application designer.
%
Commonly, in fact, applications are designed with a specific infrastructure in mind,
whose assumptions leak into the application logic.

@@

In this work,
we continue the research on pulverised systems by providing two key contributions:
first, we extend the theoretical framework of pulverisation by adding support for
\emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
the providing a practical implementation called \ourframework{} which,
to the best of our knowledge,
is the first \emph{practical framework supporting pulverisation} for a general-purpose language.
%
We exercise \ourframework{} with two case studies:
a small-scale real-wold demonstration, and a city-scale simulated scenario.

\begin{itemize}
    \item need for deployment-independent specs
    \item need for runtime reconfiguration (achieve QoS in face of changing conditions --- green computing?)
    \item declarativity over imperative approaches (note: can be compared with the trend in general softeng/build systems)
\end{itemize}

+ related work

@@

For instance, let us assume a minimal logical system counting two devices:
two rain gauges logging the water level for future reference,
and opening a valve when the water level crosses a threshold in both devices.
%
Once pulverised, the logical system would be split in ten deployable pulverised components
(that we indicate with their component symbol and an index, e.g., $\sigma_1$ is the logical sensor of the first device).
%
These pulverised components can be deployed in arbitrary hosts,
as far as they have the \emph{capabilities} required to host them
(for instance, the component hosting a rain gauge logical sensor must have an actual sensor,
while the component hosting the behaviour must be sufficiently powerful to execute the logic).
%
Thus, with \emph{no change to the application logic},
the system could be deployed on very different systems:
let us assume, for instance,
that our actual target infrastructure is not composed of two connected devices,
but it is an existing \ac{iot} system
where the rain gauges are connected to two LoRaWAN motes
(which are too weak and energy-critical to execute the logic),
the valves are controlled by a dedicated device,
the levels must be logged on a database hosted on a cloud server reachable via HTTPS,
and we have an internal edge server that we can use as we please.
%
In this case, we would deploy $\sigma_1$ and $\sigma_2$ on the motes,
$\alpha_1$ and $\alpha_2$ on the valve controller,
$\kappa_1$ and $\kappa_2$ on the cloud server,
and the remaining components on the edge server.

@@

@@

@@

In this section,
we exercise the proposed framework with two case studies.
%
In the former,
we show a small-scale demonstrator of a safety system warning people when a room capacity is exceeded
to showcase the integration of the framework with established technologies\footnote{
    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-crowd-room}
};
in the latter,
we simulate a city-scale distributed application deployed for a urban event
and show how the dynamic reconfiguration may help achieving balance between energy consumption and QoS\footnote{
    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-evaluation}
}.
%
Both experiments have been released with a permissive open-source license and have been archived for future reference on Zenodo~\cite{TBD}.

>>>>

\acrodef{dsl}[DSL]{domain-specific language}
\acrodef{iot}[IoT]{Internet of Things}
\acrodef{cas}[CAS]{collective adaptive system}
\acrodef{qos}[QoS]{quality of service}

@@

Recent research trends in topics related to the Internet of Things promote the idea of a cloud-edge continuum~\cite{DBLP:journals/iot/BittencourtISFM18},
an infrastructure composed of heterogeneous devices ranging from large and powerful cloud servers to small connected things
on which software computes in a distributed fashion,
optimising for performance and resource utilisation.
%
This kind of infrastructure is ideal for the execution of \ac{cas},
which have components that,
depending con the conditions at hand,
could benefit from being deployed on different devices.
%
Exploiting this intriguing infrastructure, however, poses new challenges to the application designer.
%
Commonly, in fact, applications are designed with a specific infrastructure in mind,
whose assumptions unavoidably leak into the application logic~\cite{Spolsky2004},
unless captured and encapsulated away from it:
the higher the coupling between the application and the infrastructure,
the harder it is for the former to exploit the latter's full potential and adapt to changes.

@@

In this work,
we continue the research on pulverised systems by providing two key contributions:
\textbf{first}, we extend the theoretical framework of pulverisation by adding support for
\emph{runtime configuration rules}, allowing pulverised components to move at runtime across different hosts;
\textbf{second}, we provide a practical implementation called \ourframework{} which,
to the best of our knowledge,
is the first \emph{practical framework supporting pulverisation} for a general-purpose language.
%
We exercise \ourframework{} with two case studies:
a small-scale real-wold demonstration, and a city-scale simulated scenario.

\begin{itemize}
    \item need for deployment-independent specs
    \item need for runtime reconfiguration (achieve \ac{qos} in face of changing conditions --- green computing?)
    \item declarativity over imperative approaches (note: can be compared with the trend in general softeng/build systems)
\end{itemize}

@@

For instance, let us assume a minimal logical system counting two devices:
two rain gauges logging the water level for future reference
and opening a valve when the water level crosses a threshold in both devices.
%
Once pulverised, the logical system would be split in ten deployable pulverised components
(that we indicate with their component symbol and an index, e.g., $\sigma_1$ is the logical sensor of the first device).
%
These pulverised components can be deployed in arbitrary hosts,
as far as they have the \emph{capabilities} required to host them
(for instance, the component hosting a rain gauge logical sensor must have an actual sensor,
while the component hosting the behaviour must be sufficiently powerful to execute the logic).
%
Thus, with \emph{no change to the application logic},
the system could be deployed on very different systems:
let us assume, for instance,
that our actual target infrastructure is not composed of two connected devices,
but it is an existing \ac{iot} system
where the rain gauges are connected to two LoRaWAN motes
(which are too weak and energy-critical to execute the logic),
the valves are controlled by a dedicated device,
the levels must be logged on a database hosted on a cloud server reachable via HTTPS,
and we have an internal edge server that we can use as we please.
%
In this case, we would deploy $\sigma_1$ and $\sigma_2$ on the motes,
$\alpha_1$ and $\alpha_2$ on the valve controller,
$\kappa_1$ and $\kappa_2$ on the cloud server,
and the remaining components on the edge server.

@@

Although the idea behind pulverisation is simple,
finding a way to implement it effectively requires tackling several challenges at different levels:
\begin{itemize}
    \item \textbf{communication}: spliting logical devices into small deployment units implies communication among them,
    thus requiring to consider networking at two levels:
    across pulverised components (intra-device) and among logical devices (inter-device);
    \item \textbf{portability}: hosts will have very diverse hardware specifications, operang systems, and software stacks,
    \item \textbf{runtime}: runtime reconf how to handle the communication between pulverised components across different runtimes?
    \item \textbf{language}: for the idea to be exploitable by designers is quintessential that
    the pulverised configuration
    and the mapping of pulverised components to the underlying infrastructure
    can be expressed nimbly and, possibly, declaratively.
\end{itemize}

@@

In principle, there are several possibilities to implement a pulverisation configurator and runtime:
as a library
(configured and executed by a target programming language),
as a framework/runtime (configured by means of configuration files and with its own runtime),
or as a \ac{dsl} (configured with its own custom syntax).
%
The approaches have different trade-offs.
%
A (well designed) library can be flexible and easier to adopt
as it does not require to learn new languages;
but it will be usable only in the language/runtime it has been designed for,
and will require its own framework/runtime to be executed
(or appropriate bindings for existing ones).
%
A framework/runtime configured via declarative files enforces declarativity at the configuration level,
but it usually dictates network protocols and features limited flexibility,
as options that have not been accounted for at design time can hardly be injected through configuration files.
%
Finally, a \ac{dsl} can be designed to be as expressive as needed,
but has high maintenance cost and, potentially, a steep learning curve due to the need of learning a new (custom) language.
%

TODO: recently, internal DSLs

@@

In this section,
we exercise the proposed framework with two case studies.
%
In the former,
we show a small-scale demonstrator of a safety system warning people when a room capacity is exceeded
to showcase the integration of the framework with established technologies\footnote{
    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-crowd-room}
};
in the latter,
we simulate a city-scale distributed application deployed for a urban event
and show how the dynamic reconfiguration may help achieving balance between energy consumption and \ac{qos}\footnote{
    \url{https://github.com/nicolasfara/ACSOS-2023-pulverization-evaluation}
}.
%
Both experiments have been released with a permissive open-source license and have been archived for future reference on Zenodo~\cite{TBD}.
